Two-Dimensional Vesicle Hydrodynamics from Hydrophobic Attraction Potential

### 2021-09-21 doc/main_jfm.tex
edited up to 4.2.4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tank-Treading Vesicles}
\begin{figure}
\begin{center}
\includegraphics[width=13.5cm]{Figure4_Wrapper.pdf}
%\hspace{-0.6cm}
%\includegraphics[height=2in]{ReducedArea.pdf}
%\hspace{0.6cm}
%\includegraphics[height=2in]{DecayRate.eps}\\
%\includegraphics[height=2in]{ArcLength.pdf}
%\includegraphics[height=2in]{ExcLength.pdf}
\end{center} 
  \caption{\label{figure4}
 JP vesicles loose enclosed area and conserve length in shear flow. 
  Panel (a) gives the reduced area over time
  for four different shear rates. The smooth curves are a best-fit
  exponential curve. There is commensurate increase in excess length (panel (b)), 
  but the arc length of the mid plane curve is more or less constant.   
   The legend in panel (b) applies to panels (a)--(c).
  Panel (d) plots the steady-state reduced area and the decay rate $k$ (inset) using the exponential fit
  to the data in panel (a). }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vesicle in a Shear Flow}
\label{sec:ves_in_shear}
Our simulation studies begin by showing, for the first time, that a JP suspension 
behaves as a tank-treading vesicle
with zero viscosity contrast between the fluids inside and
outside of the vesicle (\cite{Finken08, Shaqfeh11}).
The centroid of the baseline $58$-body JP vesicle lies at the origin 
and the background shear flow 
\begin{align}
  \uu_{\infty}(\xx) = \dot\gamma (\xx \cdot \mathbf{e}_y) \mathbf{e}_x,
\end{align} 
is applied for
shear rate $\dot\gamma$, and orthogonal unit vectors $\mathbf{e}_x$
and $\mathbf{e}_y$ for the horizontal and vertical directions, respectively.
We use the dimensionless shear rate $\chi = \dot \gamma$ ns.

@@

We point out that the range of values for $\chi$ where we measured for tank-treading 
correspond to shear rates $\dot \gamma = O(10^6$ s$^{-1}$) 
which gives fluid velocities like m s$^{-1}$ in the vicinity of the  vesicle.  
While seemingly large, these values are consistent with prior MD studies (\cite{Brandner2019})
and are a consequence of the fact that the magnitude of elastic membrane stresses grows inversely 
with the vesicle radius.  Larger shear rates are required to produce the viscous stresses needed to 
appreciably deform smaller vesicles.

\begin{figure}
\begin{center}
\includegraphics[width=11cm]{Figure5_Wrapper.pdf}\\
\includegraphics[width=11cm]{Figure5B_Wrapper.pdf}
%\includegraphics[width=0.5\textwidth]{Slip.eps}\\
%\includegraphics[width=0.4\textwidth]{Slip2.eps}
\end{center} 
  \caption{\label{figure5} The schematic of the inter-monolayer slip over
  time. Taking a shear rate $\chi=0.005$ as an example, we denote the
  mean angle between each particle pair $\theta(t)$ where the angle is
  calculated from the relative position of the particles with respect to
  the JP vesicle's center of mass. The configuration plots have two
  fixed particles marked in yellow and blue and labeled with the
  corresponding time $t=2000$ and $t=9000$. The midplane curve is
  denoted as $C_M$. We see that the inter-monolayer slip is largest
  for large shear rates.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Inter-Monolayer Friction}

The granulated JP vesicle allows us to consider inter-monolayer slip
between the two leaflets. When the shear rate is sufficiently large,
take $\chi = 0.005$ as an example, Figure~\ref{figure5} shows slip
between a pair of particles over time. The pair we consider are colored
in blue and yellow. In this series of tests, we compute the mean
particle position and mean angle $\theta(t)$ between the particle
relative positions with respect to the vesicle center. It is clear that
the outer leaflet has larger tangential velocity than the inner leaflet
since $\theta(t)$ increases. This inter-leaflet sliding phenomenon is
consistent with what has been seen in articles in molecular dynamics
simulations. (\cite{Zgorski2019, denOtter2007})

To calculate the inter-monolayer friction coefficient for the apposing
JP leaflets, we let
\begin{align}
  b = \frac{\langle F\rangle}{\langle L \rangle \langle U\rangle},
\end{align}
%
where 
\begin{align}
  \langle F\rangle = \left\langle \int_{\mathcal{C}_M} 
    \jump{\ttau \cdot \sigma \cdot \nnu} \,\dif s \right\rangle
\end{align}
is the time-averaged, tangential force along the midplane curve
$\mathcal{C}_M$. The parameters $\langle L\rangle$ and $\langle
U\rangle$ are the time-averaged midplane curve length and intermonolayer
slip velocity, respectively.

We calculate $\langle F\rangle$ by first evaluating the smooth
tangential force along curves a positive distance inside and outside
$\mathcal{C}_M$ respectively, and then interpolating at zero distance to
give a force jump. We have made a slight abuse of language: due to the
problem being two-dimensional, the integral $\langle F\rangle$ is
actually a force density per length which when divided by length gives
the requisite force per area. 
%which when divided by length

Table~\ref{table1} provides dimensional friction coefficients $b$ for
several shear rates and the results are in quantitative agreement with
values previously reported in the literature (\cite{sch-vla-mik2010}
and~\cite{denOtter2007}). The coefficients generally lie within the
range 0.2 to 0.3 pN$\cdot$ns nm$^{-3}$ but there are some outliers.
These outliers ($\dot\gamma = 0.0041$ ns$^{-1}$ and $\dot\gamma =
0.0066$ ns$^{-1}$) correspond to slip velocities that are too low
relative to the shear rate (see the green and red curve in
Figure~\ref{figure5}(c)). The JP vesicle consists of only 58 particles
and so it is not surprising that discrete effects produce outliers.

%\todo[inline]{If $B$ is the numerically calculated friction coefficient,
%then $b = B/2.5^3$ pN ns nm$^{-3}$ gives the dimensional value.
%
%If $P$ is the numerically calculated pressure, then $p = P/2.5$ pN
%nm$^{-2}$ is the dimensional pressure.}

\begin{table}
\caption{Friction Coefficients}
\centering
\begin{tabular}{c c c c c c c c }
 $\chi$ & 0.0020   &  0.0025 &  0.0030 &  0.0035 &  0.0040 & 0.0045 & 0.0050  \\
\hline                    
$\dot\gamma$ (ns$^{-1}$)        & 0.0033 & 0.0041 & 0.0049 & 0.0057 & 0.0066 & 0.0074 & 0.0082\\
%$F_{f}$ (pN nm$^{-1}$)           & 1.84 & 4.90 & 11.36 & 8.16 & 6.14 &  17.69 &  24.33 \\[1ex]
%$\dot\theta$ (nm ns$^{-1}$)           & $6.84\times10^{-5}$ & $1.31\times10^{-4}$ & $3.46\times10^{-4}$ & $2.32\times10^{-4}$ & $1.81\times10^{-4}$ & $4.91\times10^{-4}$ & $5.83\times10^{-4}$\\ [1ex]
$b$ (pN$\cdot$ns nm$^{-3}$)    & 0.22 & 0.87  & 0.22 &0.21 & 1.29 & 0.31 & 0.30\\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}

\subsubsection{Membrane Ruptures}
\begin{figure}
\centering
\includegraphics[width=11.32cm]{Figure8_Wrapper.pdf}
%\includegraphics[height=2in]{N58_rupt_0.pdf}
%\includegraphics[height=2in]{N58_rupt_200.pdf}
\\
%\includegraphics[height=2in]{N58_rupt_400.pdf}
%\includegraphics[height=2in]{N58_rupt_600.pdf}
  \caption{\label{figure8} Membrane rupture occurs when the shear rate
  exceeds a threshold. Here the shear rate is $\chi = 0.1$. Starting
  with a circular shape (panel (a)), the vesicle is stretched by the
  background flow and multiple pores form (panel (b)). Panels (c)--(d)
  show the later states during the simulation. Streamlines are included
  in each panel.}
\end{figure}
A pore formation or a complete membrane rupture can occur at large shear
rates. Figure~\ref{figure8} demonstrates how a vesicle can rupture when
suspended in a shear flow with shear rate $\chi=0.1$. The streamlines
are included in these plots. An interesting phenomenon occurs in panels
(c) and (d) where two separated bilayers move along the background shear
flow symmetrically. Consequently, a periodic motion for bilayers may
occur at certain shear rates. One can expect that these two layers will
diverge from one another for much higher shear rates.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tank-Treading Vesicles}
\begin{figure}
\begin{center}
\includegraphics[width=13.5cm]{Figure4_Wrapper.pdf}
%\hspace{-0.6cm}
%\includegraphics[height=2in]{ReducedArea.pdf}
%\hspace{0.6cm}
%\includegraphics[height=2in]{DecayRate.eps}\\
%\includegraphics[height=2in]{ArcLength.pdf}
%\includegraphics[height=2in]{ExcLength.pdf}
\end{center} 
  \caption{\label{figure4}
 JP vesicles loose enclosed area and conserve length in shear flow. 
  Panel (a) gives the reduced area over time
  for four different shear rates. The smooth curves are a best-fit
  exponential curve. There is commensurate increase in excess length (panel (b)), 
  but the arc length of the mid plane curve is more or less constant.   
   The legend in panel (b) applies to panels (a)--(c).
  Panel (d) plots the steady-state reduced area and the decay rate $k$ (inset) using the exponential fit
  to the data in panel (a). }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vesicle in a Shear Flow}
\label{sec:ves_in_shear}
Our simulation studies begin by showing, for the first time, that a JP suspension with hydrophobic attraction 
behaves as a tank-treading vesicle
with zero viscosity contrast between the fluids inside and
outside of the vesicle (\cite{Finken08, Shaqfeh11}).
The centroid of the baseline $58$-body JP vesicle lies at the origin 
and the background shear flow 
\begin{align}
  \uu_{\infty}(\xx) = \dot\gamma (\xx \cdot \mathbf{e}_y) \mathbf{e}_x,
\end{align} 
is applied for
shear rate $\dot\gamma$, and orthogonal unit vectors $\mathbf{e}_x$
and $\mathbf{e}_y$ for the horizontal and vertical directions, respectively.
We use the dimensionless shear rate $\chi = \dot \gamma$ ns.

\begin{figure}
\begin{center}
\includegraphics[width=11cm]{Figure5_Wrapper.pdf}\\
\includegraphics[width=11cm]{Figure5B_Wrapper.pdf}
%\includegraphics[width=0.5\textwidth]{Slip.eps}\\
%\includegraphics[width=0.4\textwidth]{Slip2.eps}
\end{center} 
  \caption{\label{figure5} 
  There is inter-monolayer slip present in the JP vesicle tank-treading motion at all shear rates.  
  Panel (a) tracks the angle  $\theta(t)$ between a pair of particles (blue and yellow) with respect to
  the JP vesicle centroid. The shear rate is $\chi=0.005$ and in the right figure, the yellow particle
  will complete two and quarter revolutions in the same time the blue particle completes two revolutions.
  The curves in panel (b) are the mean change in angle for various shear rates. 
  Slip velocity generally increases with shear rate but 
  we have marked two outliers (times symbols). }
\end{figure}

@@

We point out that the range of values for $\chi$ where we measured for tank-treading 
correspond to shear rates $\dot \gamma = O(10^6$ s$^{-1}$) 
which gives fluid velocities like m s$^{-1}$ in the vicinity of the  vesicle.  
While seemingly large, these values are consistent with prior MD studies (\cite{Brandner2019})
and are a consequence of the fact that the magnitude of elastic membrane stresses grows inversely 
with the vesicle radius; larger shear rates are required to produce the viscous stresses needed to 
appreciably deform smaller vesicles.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Inter-Monolayer Friction}
We observe inter-monolayer slip in the tank-treading, JP vesicle at all shear rates. 
The two leaflets of the vesicle are not bound together, but rather are able to slide 
past one another. This sliding is limited by viscous friction of the aqueous gaps between particles
and by the constant unbinding and binding of particles pairs in apposing leaflets.  
The phenomenon is consistent with prior MD results (\cite{Zgorski2019, denOtter2007}).

Figure~\ref{figure5}a illustrates the process by tracking the angle $\theta(t)$
between two particles, one blue and the other yellow, relative to the vesicle centroid. 
It is clear from the growth in angle that the outer leaflet has a tangential velocity that is larger  
then that of the inner leaflet.  
Figure~\ref{figure5}b plots the mean, particle-pair angle for various shear rates. 
It suggests that the inter-monolayer slip velocity generally increases with shear rate.

From the data, we are able to define an inter-monolayer friction coefficient  
\begin{align}
  b = \frac{\langle F\rangle}{\langle L \rangle \langle U\rangle},
\end{align}
where $F$ is the tangential force jump, $L$ is the length of the midplane $C_M$ (Figure~\ref{figure5}a), $U$ is the slip velocity, and 
$\langle \cdot \rangle$ denotes the time average. 
To measure inter-monolayer slip $U$, we compute the mean tangential velocity of the particles in the outer monolayer
and subtract from it the mean tangential velocity of the particles in the inner monolayer. 
To calculate the tangential force jump $F$, we let 
\begin{equation}
F_h = \int_{\mathcal{C}_M^+} 
    \ttau \cdot \sigma \cdot \nnu \,\dif s -  \int_{\mathcal{C}_M^-} 
    \ttau \cdot \sigma \cdot \nnu \,\dif s 
\end{equation}
where $C_M^+$ and $C_M^-$ are target curves obtained by moving the midplane curve $C_M$ (Figure~\ref{figure5}a)
a distance $h$ along the outward, respectively inward, unit normal.  Actually, we calculate $F_h$ for several values of $h$ 
and then define $F$ to by extrapolating to zero distance giving a jump in tangential shear force across the midplane.

\begin{table}
\caption{Friction Coefficients}
\centering
\begin{tabular}{c c c c c c c c }
 $\chi$ & 0.0020   &  0.0025 &  0.0030 &  0.0035 &  0.0040 & 0.0045 & 0.0050  \\
\hline                    
$\dot\gamma$ (ns$^{-1}$)        & 0.0033 & 0.0041 & 0.0049 & 0.0057 & 0.0066 & 0.0074 & 0.0082\\
%$F_{f}$ (pN nm$^{-1}$)           & 1.84 & 4.90 & 11.36 & 8.16 & 6.14 &  17.69 &  24.33 \\[1ex]
%$\dot\theta$ (nm ns$^{-1}$)           & $6.84\times10^{-5}$ & $1.31\times10^{-4}$ & $3.46\times10^{-4}$ & $2.32\times10^{-4}$ & $1.81\times10^{-4}$ & $4.91\times10^{-4}$ & $5.83\times10^{-4}$\\ [1ex]
$b$ (pN$\cdot$ns nm$^{-3}$)    & 0.22 & 0.87*  & 0.22 &0.21 & 1.29* & 0.31 & 0.30\\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}

Table~\ref{table1} provides dimensional friction coefficients $b$ for
several shear rates. The coefficients generally lie within the
range 0.2 to 0.3 pN ns nm$^{-3}$ which is in good quantitative agreement with
values previously reported in the literature (\cite{sch-vla-mik2010}
and~\cite{denOtter2007}).  
The data contain two outliers (Table~\ref{table1}, * symbol).
These correspond to the two outliers in the slip velocities (Figure~\ref{figure5}b, times symbol). 
We suspect that due to the finite particle size, the granular wobble of the tank-treading motion is more prominent for some
shear rates, causing some slip velocities to be lower than expected.

\subsubsection{Membrane Ruptures}
\begin{figure}
\centering
\includegraphics[width=11.32cm]{Figure8_Wrapper.pdf}
%\includegraphics[height=2in]{N58_rupt_0.pdf}
%\includegraphics[height=2in]{N58_rupt_200.pdf}
\\
%\includegraphics[height=2in]{N58_rupt_400.pdf}
%\includegraphics[height=2in]{N58_rupt_600.pdf}
  \caption{\label{figure8} 
  Tank-treading leads to membrane rupture for large shear rates.
  }
\end{figure}
A temporary fissure or a complete membrane rupture can occur at large shear
rates. Figure~\ref{figure8} demonstrates how a vesicle can rupture when
suspended in a shear flow. For $\chi = 0.1$, starting
  with a circular shape (Figure~\ref{figure8}a), the vesicle is stretched by the
  background flow and fissures appear in the bilayer structure in multiple locations (Figure~\ref{figure8}b). 
  In Figure~\ref{figure8}c, the ruptured vesicles forms two planar micelles which 
  are eventually carried off by the flow (Figure~\ref{figure8}d).
